Sharing statistics for SPARQL Federation optimization, with emphasis on benchmark quality

### 2012-01-17 eswc-2012/phdsymp.tex
siste Audun-rettinger
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@103 43ef1190-90fd-4dd7-a769-d0121f02b80e

Query federation with SPARQL, which is a standardized query language
for the Semantic Web, has attracted much attention
from industry and academia alike, and four implementations of basic
query federation were submitted to the SPARQL 1.1 Working Group as
input for the forthcoming
work\footnote{http://www.w3.org/2009/sparql/wiki/Feature:BasicFederatedQuery}. 
This feature was
supported by a many group members, and the Last Call working
draft of the proposed standard was published on 17~November~2011.

@@

This problem has already received some attention, in \cite{5337556} it
is shown how various statistics-gathering techniques, including
histograms, can be computed to assist the SemWIQ optimizer, written by
the same authors.  However, histogram approaches generally suffer
from the problem that they grow too large or become a insufficiently
accurate digest, especially in the face of very heterogeneous data.
\cite{Harth:2010:DSO:1772690.1772733} introduced QTrees, which may
alleviate the problem of histogram size, but which may not solve it.

@@

With this analysis, I speculate based on superficial experience that
certain estimated coefficients can be exposed in the service
description to give federated query engines what they need to optimize
for performance characteristics of certain SPARQL
implementations. Adjusting how different features are weighed will
help those who know details of their expected SPARQL use and
normalization-techniques of the weights will answer the special case
of ``the best general-purpose SPARQL engine''.

>>>>

Query federation with SPARQL, which is a standardized query language
for the Semantic Web, has attracted much attention
from industry and academia alike, and four implementations of basic
query federation were submitted to the SPARQL 1.1 Working Group as
input for the forthcoming
work\footnote{http://www.w3.org/2009/sparql/wiki/Feature:BasicFederatedQuery}. 
This feature was
supported by many group members, and the Last Call working
draft of the proposed standard was published on 17~November~2011.

@@

This problem has already received some attention, in \cite{5337556} it
is shown how various statistics-gathering techniques, including
histograms, can be computed to assist the SemWIQ optimizer, written by
the same authors.  However, histogram approaches generally suffer
from the problem that they grow too large or become an insufficiently
accurate digest, especially in the face of very heterogeneous data.
\cite{Harth:2010:DSO:1772690.1772733} introduced QTrees, which may
alleviate the problem of histogram size, but which may not solve it.

@@

With this analysis, I speculate based on superficial experience with
factorial designs and analysis of variance that certain estimated
coefficients can be exposed in the service description to give
federated query engines assistance in optimizing for performance
characteristics of certain SPARQL implementations. Adjusting how
different features are weighed will help those who know details of
their expected SPARQL use and normalization-techniques of the weights
will answer the special case of ``the best general-purpose SPARQL
engine''.
